---
title: "Viz Part 1"
output: github_document
date: "2023-10-03"
---

```{r}
library(tidyverse)
library(ggridges)
```

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

Get the data for plotting today

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


Let's make a plot!

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

Pipes and stuff. Because it's tidyverse, you can take whatever dataset you have and then pipe it into ggplot and define the aesthetics, as long as the first thing you are putting in is a dataframe.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_nyc_weather = 
  weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_nyc_weather
```


##Fancy plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3) +
  geom_smooth(se = FALSE)
```


Plot with facets

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  facet_grid(. ~ name)
```


Trying again on a different plot.

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .3) + 
  geom_smooth() +
  facet_grid(. ~ name)
```

assigning specific colors. Almost never assign colors by hand. Ggplot assigns some colors by default based on how many options you have.

```{r}
weather_df |> 
  filter(name != "CentralPark_NY") |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = .7, size = .5)
```

hex plot :-)
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex(alpa = .3)
```


## univariate plotting, for when you just want to look at one variable at a time

histogram.... it is ALWAYS the SAME!! GGplot.. then dataframe.. then what variable maps onto what axis... then what kind of graph you want !

avoid stacking up bars in histrogram with "dodge"

```{r}
ggplot(weather_df, aes(x= tmax, fill = name)) +
  geom_histogram(position = "dodge")
```

let;s use a density plot
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_density(alpha = .3, adjust = .5)
```

using boxplots
```{r}
ggplot(weather_df, aes(y = tmax, x = name)) +
  geom_boxplot()
```


violin plots

```{r}
ggplot(weather_df, aes(y = tmax, x = name)) +
  geom_violin()
```

ridge plot... I feel like this is on NYTimes a lot ...

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) +
  geom_density_ridges()
```


## saving and embedding plots



```{r}
ggplot_weather = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax))
```

ggplot_weather



ggsave("results/ggplotweather.pdf", ggplot_weather)
